---
title: "Haplotype distribution"
author: "Reuben M. Yaa"
date: "01/11/2022"
output: html_document
---

```{r setup, include=FALSE}
root.dir <- here::here()
knitr::opts_chunk$set(echo = TRUE, root.dir=root.dir)
knitr::opts_knit$set(root.dir = root.dir)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Introduction
Assessed the distribution density of the haplotypes through out the infection episodes and mapped back the haplotypes to the patients and their respective infection episodes occurring in highest distribution. The pattern of the top 4 haplotypes i.e E-KSNG-L, Q-KSNG-F, Q-KSNG-L and Q-KNNG-L was followed

## Haplotype distribution
```{r results = FALSE}
library("tidyverse")
#library("gghighlight")
library(scales)
hap_den<- readxl::read_xlsx("data/Haplotype Distribution.xlsx", col_names = TRUE)
hap_den<-rbind(hap_den)

dat <- with(density(hap_den$Episodes), data.frame(x, y))
infec<-ggplot(data = dat, mapping = aes(x = x, y = y), lwd = 4) +
  labs(x="infection episodes", y = "Density")+
  scale_x_continuous(labels = scales::comma)+
    theme_classic(base_size = 10) +
  theme(aspect.ratio=5/7)+theme(text = element_text(size = 20))+
    geom_line()+
    geom_area(mapping = aes(x = ifelse(x>2 & x< 15 , x, 0)), fill = "grey") +
    xlim(1, 30)
infec
pdf("figures/infection_distribution.pdf", width = 15, height = 5, pointsize=100)#save as pdf file
infec
dev.off()
```

## Haplotype frequencies
Following the frequency of top 4 haplotypes i.e E-KSNG-L, Q-KSNG-F, Q-KSNG-L and Q-KNNG-L at the peak haplotype distribution
```{r results = FALSE}
library("tidyverse")

freq<-ggplot(hap_den[hap_den$Episodes>1 & hap_den$Episodes< 16 ,], aes(y= Frequency, x=Episodes)) +
  stat_smooth(aes(group= Haplotype, color= Haplotype, fill=Haplotype), method="loess", size=1, formula="y ~ log(x)", se=FALSE) +
  geom_point(aes(colour=Haplotype))+
  labs(x="Infection episodes", y = "Frequency")+ facet_grid(Haplotype ~ .)+
  theme_classic(base_size = 10) +
  theme(aspect.ratio=5/9)+theme(text = element_text(size = 20))
   
freq
pdf("figures/haplotype_distribution.pdf", width = 15, height = 10, pointsize=30)#save as pdf file
freq
dev.off()

```

Figure available [haplotype distribution](https://mangiruben.github.io/pfMSP1-Malaria/figures/haplotype_distribution.pdf)
<hr>