---
title: "Microhaplotype distribution"
author: "Reuben M. Yaa"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output: html_document
---

```{r setup, include=FALSE}
root.dir <- here::here()
knitr::opts_chunk$set(echo = TRUE, root.dir=root.dir)
knitr::opts_knit$set(root.dir = root.dir)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Introduction
Assessed the distribution density of the microhaplotypes through out the infection episodes and mapped back the microhaplotypes to the patients and their respective infection episodes occurring in highest distribution. The pattern of the top 4 microhaplotypes i.e E-KSNG-L, Q-KSNG-F, Q-KSNG-L and Q-KNNG-L was followed


## Microhaplotype frequencies
Following the frequency of top 4 microhaplotypes i.e E-KSNG-L, Q-KSNG-F, Q-KSNG-L and Q-KNNG-L at the peak haplotype distribution

```{r results = FALSE}
library("tidyverse")
#library("gghighlight")
library(scales)
hap_den<- readxl::read_xlsx("data/Haplotype Distribution.xlsx", col_names = TRUE)
hap_den<-rbind(hap_den)
hap_den<-as.data.frame(hap_den)

freq <- ggplot(hap_den[hap_den$Episodes > 1 & hap_den$Episodes < 16, ], aes(y = Frequency/33, x = Episodes)) +
  geom_line(aes(group = Haplotype), color = "black") +
  geom_point(color = "black") +  # Change lines and dots to black
  labs(x = "Infection episodes", y = "Frequency") +
  facet_wrap(~ Haplotype, scales = "free_y", nrow = 1) +
  theme_classic(base_size = 10) +
  theme(aspect.ratio = 5 / 7, text = element_text(size = 20)) +
  ylim(0, 0.2)

freq

pdf("figures/haplotype_distribution.pdf", width = 10, height = 7)#save as pdf file
freq
dev.off()

```

## Microhaplotype distribution
```{r results = FALSE}
hap_den$Haplotype <- ifelse(hap_den$Haplotype %in% c("E-KSNG-L", "Q-KSNG-L", "Q-KSNG-F", "Q-KNNG-L"), "Prevalent", hap_den$Haplotype)


dat <- with(density(hap_den$Episodes), data.frame(x, y))
infec<-ggplot(data = dat, mapping = aes(x = x, y = y), fill=Haplotype, colour=Haplotype, lwd = 4) +
  labs(x="infection episodes", y = "Density")+
  scale_x_continuous(labels = scales::comma)+
    theme_classic(base_size = 15) +
  theme(aspect.ratio=5/7)+theme(text = element_text(size = 20))+
    geom_line()+
    geom_area(mapping = aes(x = ifelse(x>2 & x< 15 , x, 0)), fill = "grey") +
    xlim(1, 30)
infec
pdf("figures/infection_distribution.pdf", width = 15, height = 5, pointsize=100)#save as pdf file
infec
dev.off()
#classifying the density distributions as either prevalent or rare
infec_2<-ggplot(hap_den, aes(x=Episodes, fill=Haplotype, colour=Haplotype )) +
  geom_density(alpha=0.4, adjust = 1.25, lwd = 0.5) +
    labs(x="infection episodes", y = "Density")+
  scale_x_continuous(labels = scales::comma)+
theme_classic(base_size = 15)+
theme(aspect.ratio=5/7)+theme(text = element_text(size = 25))+
  xlim(1, 30)
infec_2
pdf("figures/prevalent_rare_distribution.pdf", width = 15, height = 5, pointsize=16)#save as pdf file
infec_2
dev.off()
```



Figure available [microhaplotype distribution](https://mangiruben.github.io/pfMSP1-Malaria/figures/haplotype_distribution.pdf) and [microhaplotype class distribution](https://mangiruben.github.io/pfMSP1-Malaria/figures/prevalent_rare_distribution.pdf)
<hr>