---
title: "Microhaplotypes per patient"
author: "Reuben M. Yaa"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output: html_document
---

```{r setup, include=TRUE}
root.dir <- here::here()
knitr::opts_chunk$set(echo = TRUE, root.dir=root.dir)
knitr::opts_knit$set(root.dir = root.dir)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Introduction
Assigned back the microhaplotypes respective patients irrespective of the infection episodes. This identifies the patients that were mostly infected and by which haplotype. 

### Patients microhaplotypes distribution
```{r}
library(plotly)
library(dplyr)
library(networkD3)
library(ggplot2)

hap<- readxl::read_xlsx("data/Haplotype assigned.xlsx", col_names = TRUE)
hap <-as.data.frame(hap)

geno<- readxl::read_xlsx("data/genotype+props.xlsx", col_names = TRUE)
geno <-as.data.frame(geno)
sankey <- function(data, var1, var2, count = "n"){
## Rename
  data$pre_group <- data[[var1]]
  data$group <- data[[var2]]
  ## binding nodes to one object
  group_source <- unique(data$pre_group)
  group_target <- paste0(unique(data$group), " ")
  groups <- c(group_source, group_target)
  nodes_source <- tibble(name = group_source)
  nodes_target <- tibble(name = group_target)
  nodes <- rbind(nodes_source, nodes_target) %>% mutate(node = 0:(nrow(.) - 1))
  
  ## Set up `links`
  links <-
    data %>%
    mutate(group = paste0(group, " ")) %>%
    select(source = "pre_group",
           target = "group",
           value = count)
  
  ## setting up data frame from columns
  nodes_source <- nodes_source %>% select(name) 
  nodes_target <- nodes_target %>% select(name) 
  
  links <-
    links %>%
    mutate(source = as.character(source)) %>% 
    left_join(nodes %>% rename(IDsource = "node"), by = c("source" = "name")) %>%
    left_join(nodes %>% rename(IDtarget = "node"), by = c("target" = "name"))
  
  networkD3::sankeyNetwork(Links = as.data.frame(links),
                           Nodes = as.data.frame(nodes),
                           Source = 'IDsource', 
                           Target = 'IDtarget', 
                           Value = 'value',
                           NodeID = 'name',
                           units="count",
                           fontSize = 16,
                           nodePadding = 7,
                           nodeWidth = 20,
                           sinksRight = FALSE)
}

sankey<-sankey(hap, var1="Patient_ID", var2="Haplotype", count = "Frequency")
sankey
```

### prevalent microhaplotype proportion 
Reviewed to include the proportion of the prevalent prevalent haplotype  in relation to all the microhaplotypes per patient
```{r}
library(productplots)
library(ggplot2)
#hap['Haplotype'][hap['Haplotype']!="E-KSNG-L"]<-"OTHER"


hap$Haplotype[hap$Haplotype !="E-KSNG-L" &  hap$Haplotype != "Q-KSNG-L"  & hap$Haplotype != "Q-KSNG-F" & hap$Haplotype != "Q-KNNG-L"] <-"OTHER"

order =c(23, 30,7,8,16,19,20,29,1,2,10,14,21,27,28,9,11,13,18,26,3,12,32,15,25,17,  24, 33,31,22)#from Hierachy of next figure
                                     
props<-ggplot() + 
geom_bar(aes(x = factor(hap$Patient_ID, levels = order), y = Frequency, fill = factor(Haplotype, levels=c("OTHER","E-KSNG-L", "Q-KSNG-L", "Q-KSNG-F", "Q-KNNG-L"))), data = hap,position="fill", stat="identity")+
    ylab("Proportion")+coord_flip()+
    labs(fill="Haplotype")+
    theme_classic() +
     #scale_x_continuous(breaks = seq(1, 33, by = 1))+
      
    theme(axis.line.y = element_blank(), axis.line.x = element_line(), text = element_text(size = 14), axis.text.x = element_text(size = 16))
    

### Saving the plot
props 
pdf("figures/props_distribution.pdf", width = 5, height = 8, pointsize = 16)#save as pdf file
props
dev.off()




geno<- readxl::read_xlsx("data/genotype+props.xlsx", col_names = TRUE)
geno <-as.data.frame(geno)

#mean_SD is mean genotyped per individual+SD
mean_SD <- c(mean(geno$Episodes[geno$state == "genotyped"]), sd(geno$Episodes[geno$state == "genotyped"]))


#geno <- geno[order(-geno$diff),]

p <- ggplot(geno, aes(y = reorder(Patient_ID, diff), x = diff)) +
  geom_col(aes(fill = factor(state, levels=c("total_epis", "genotyped"))))+

geom_text(aes(label = diff), size = 3, hjust = 1.5, position = "stack") +
  
  scale_x_continuous(breaks = seq(0, 30, by = 5))+ theme_classic() +
    theme(axis.line.y = element_blank(), axis.line.x = element_line(), text = element_text(size = 14), axis.text.x = element_text(size = 16) )+
stat_summary(fun.y = sum, aes(label = ..x..), geom = "text", hjust = -0.3, size = 0)+#changed size to 0, suggestions to remove the total(genotyped+non_genptyped)
  scale_fill_manual(values=c('#DCDCDC','#B6B6B4' ))

p + labs(fill = "state")
pdf("figures/genotyped_distribution.pdf", width = 5, height = 8, pointsize = 16)
p + labs(fill = "state")
dev.off()

```



```{r, results=FALSE}
saveNetwork(sankey, "figures/haplotypes_assigned.html")
png<-webshot::webshot("figures/haplotypes_assigned.html","figures/haplotypes per patient.pdf", vwidth = 1000, vheight = 900)
unlink("figures/haplotypes_assigned*", recursive = TRUE)
```


### Patterns of consecutive infection

```{r}
sup<- readxl::read_xlsx("data/Supplementary Table 1.xlsx", col_names = TRUE)
sup <-as.data.frame(sup)


sup_filtered <- sup[sup$Patient_ID %in% unique(sup$Patient_ID[diff(sup$Episode) == 1]), ]

#Now sup_filtered contains rows where episodes are consecutive per patient

sup_filtered$Episode <- as.numeric(sup_filtered$Episode)


# Function to filter consecutive episodes with the same haplotypes within a group
consecutive_same_haplotypes_within_group <- function(patient_data) {
  consecutive_rows <- patient_data$Episode == lag(patient_data$Episode) + 1 & 
                       patient_data$Haplotype == lag(patient_data$Haplotype)
  return(patient_data[consecutive_rows | lead(consecutive_rows), ])
}


consecutive_same_haplotypes_data <- sup_filtered %>%
  group_by(Patient_ID) %>%
  do(consecutive_same_haplotypes_within_group(.))

# Ungroup the result
consecutive_same_haplotypes_data <- ungroup(consecutive_same_haplotypes_data)
consecutive_same_haplotypes_data <- consecutive_same_haplotypes_data %>%
  filter(!is.na(Patient_ID))
spp<- readxl::read_xlsx("data/hap.xlsx", col_names = TRUE)
spp <-as.data.frame(spp)

#state time between similar consecutive infections in months

spp$Patient_ID <- factor(spp$Patient_ID, levels = unique(spp$Patient_ID))

# ploting the density, but going to drop this
ggplot(spp, aes(x = Episode, y = Patient_ID, fill = state)) +
    geom_tile() +
    scale_fill_gradient(low = "grey", high = "blue", limits = c(0, 17)) +
    labs(title = "Density for similar consecutive infections",
         x = "Episode",
         y = "Patient_ID") +
    theme_minimal()

spp <- spp[spp$state != 0, ]
spp<- readxl::read_xlsx("data/spp.xlsx", col_names = TRUE)
spp<- spp[complete.cases(spp$state), ]
i_order<-c("6_7", "3_4", "4_5", "6_7", "1_2", "22_24", "23_24", "17_18","2_3","9_10", "10_11", "9_10", "13_14" )
i_order_unique <- unique(i_order)

# Convert Episode to a factor with custom levels
spp$Episode <- factor(spp$Episode, levels = i_order_unique)


pdf("figures/infect_interval.pdf", width = 4, height = 4, pointsize = 18)
ggplot(spp, aes(x = Episode, y = Patient_ID, fill = Haplotype, size = state)) +
  geom_point(shape = 21) +
  scale_size_continuous(range = c(1, 7), breaks = seq(1, 16, by = 4)) +
  labs(title = "Consecutive infec intervals",
       x = "Consecutive infections",
       y = "Patient_ID") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.ticks = element_line(color = "black"),  
    panel.border = element_rect(color = "black", fill = NA),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    axis.line = element_line(color = "black")  
  )

dev.off()
```


#### Distribution of interval of infection using histogram
```{r}
infe<- readxl::read_xlsx("data/MSP1_sample_data.xlsx", col_names = TRUE)
infe <-as.data.frame(infe)
infe <- infe[complete.cases(infe$interval_inf), ]


inf<-hist(infe$interval_inf, breaks=30, main = "Distribution of interval of infections", xlab = "Infection interval in months", ylim = c(0,100))
#inf
pdf("figures/freq_infect_interval.pdf", width = 4, height = 6, pointsize = 16)
hist(infe$interval_inf, breaks=30, main = "Distribution of interval of infections", xlab = "Infection interval in months", ylim = c(0,100))
dev.off()

```


Figure available [haplotype per patient](https://mangiruben.github.io/pfMSP1-Malaria/figures/haplotypes per patient.pdf) [proportion_distributions](https://mangiruben.github.io/pfMSP1-Malaria/figures/figures/props_distribution.pdf) and [Infection intervals in months](https://mangiruben.github.io/pfMSP1-Malaria/figures/figures/freq_infect_interval.pdf)
<hr>